### 2025-05-20 paper/discussion.tex

\section{Discussion}

\subsection{Interpretation of Results}

In our experiments we compared single-neuron and dual-neuron output layers for binary classification which reveal several consistent patterns that show deeper interpretations.

\subsubsection{Consistent Performance Advantage of Single-Neuron Approach}

The single-neuron model always consistently outperformed its dual-neuron counterpart model across all tested model architectures and datasets though the magnitude of this advantage varied slightly on some cases. We attribute this performance difference to several factors listed below:

\begin{enumerate}
\item \textbf{Optimization Landscape}: The single-neuron approach creates a simpler optimization landscape with fewer parameters and potentially allowing gradient-based optimization to find better solutions more efficiently and faster than other approach.

\item \textbf{Problem-Model Alignment}: Binary classification is inherently a one dimensional problem (decision boundary between two classes) which aligns more perfectly and naturally with the single-neuron formulation of the architecture. The dual-neuron approach introduces an extra degree of freedom that may not be necessary for the binary decision and network flow diverts with two neurons in the end which might complicate optimization landscape of this approach.

\item \textbf{Implicit Regularization}: The parameter reduction in the single-neuron approach serves as a form of implicit regularization potentially improving generalization by constraining the model's capacity which helps to mitigate overfitting issue in the models as well.
\end{enumerate}

\subsubsection{Architecture Dependency}

While the single-neuron approach outperformed across architectures the magnitude of its advantage decreased with larger more advanced like ResNet50, modern like ViT and expressive models:

\begin{enumerate}
\item \textbf{Diminishing Returns}: As model capacity increases the relative importance of output layer configuration diminishes. In high-capacity models like ResNet50 and ViT both approaches can effectively learn the decision boundary making the choice less critical in these architectures.

\item \textbf{Feature Quality}: In pre-trained networks the quality of learned features may overshadow the effect of output layer design. When the penultimate layer produces highly discriminative features the specific output layer design becomes less important and less effective in these architectures.

\item \textbf{Regularization Effects}: In larger models other regularization techniques (dropout, batch normalization, etc.) may have a more dominant effect than the implicit regularization (less parameters) provided by the single-neuron approach which helps dual neuron models to keep up with single neuron models in terms of performance.
\end{enumerate}

\subsection{Theoretical Insights}

Our experimental results across multiple architectures and classification tasks provide rich material for theoretical analysis regarding model optimization task dependent learning dynamics and the crucial role of output layer design in the model.

\subsubsection{Representational Equivalence vs. Learning Dynamics}

From a theoretical perspective both output layer approaches have equivalent representational power for binary classification. Any decision boundary that can be learned by one approach can in principle be learned by the other as well. The observed performance differences must therefore stem from differences in learning dynamics rather than representational capacity of the models:

\begin{enumerate}
\item \textbf{Parameter Coupling}: In the dual-neuron approach the softmax activation introduces coupling between two output neurons as they must sum to one (sum of all outputs must be equal to 1 to make probabilities evenly distributed). This coupling creates dependencies that can complicate the optimization landscape of the model and even worse to unstable training with late convergence.

\item \textbf{Gradient Flow}: Analysis of gradient magnitudes during training revealed (not explicitly tested rather inferred from the training curves) that the single-neuron approach tends to produce more stable and consistent gradients throughout training potentially leading to more reliable weight updates than its couterpart models.

\item \textbf{Information Bottleneck Perspective}: The single-neuron output can be viewed as creating a more severe information bottleneck which according to information bottleneck theory may lead to better generalization by forcing the model to extract only the most relevant features (this might be reason why single-neuron approach usually performed better on some tasks than dual-neuron approach).
\end{enumerate}

\subsubsection{Task Complexity and Output Layer Interaction}

A particularly illuminating finding from our experiments is the task-dependent behavior of the dual-neuron model:

\begin{enumerate}
\item \textbf{Feature Space Characteristics}: The airplane vs. automobile task (where dual-neurons performed well) likely has more linearly separable features than the cat vs. dog task (where dual-neurons performed reasonably). This suggests the dual-neuron approach may be more sensitive to the geometry (objects in the image and their representation) of the feature space.

\item \textbf{Optimization Pathology}: The complete failure of the dual-neuron model on certain tasks represents an optimization pathology that cannot be explained by only performance differences. This suggests the dual-neuron approach may be vulnerable to specific initialization conditions that lead to symmetric weights between output neurons creating a situation where gradients cancel each other or don't update the weights at all.

\item \textbf{Feature Ambiguity Benefits}: Interestingly more challenging cat vs. dog task actually helped the dual-neuron model learn. This counter intuitive result suggests that tasks with more ambiguous features may provide more varied gradient signals that help the dual-neuron model escape poor optimization regions (this might be reason why dual-neuron approach performed better on some tasks than single-neuron approach however this is theorized).
\end{enumerate}

\subsubsection{Relationship to Decision Boundary Geometry}

The structure of the output layer has implications for the geometry of the learned decision boundary:

\begin{enumerate}
\item \textbf{Direct Boundary Modeling}: The single-neuron approach with sigmoid activation function directly models the decision boundary between two classes which aligns well with fundamental nature of binary classification problem.

\item \textbf{Indirect Boundary Derivation}: The dual-neuron approach with softmax activation function derives this decision boundary indirectly from comparison of two independently modeled class probabilities potentially introducing unnecessary complexity (this complexity can gradient flow between two neurons or updating weights of one neuron based on the other).

\item \textbf{Boundary Regularization}: The reduced number of parameters of the single-neuron approach might also implicitly favor simpler decision boundaries which could explain that its better generalization especially in smaller models like custom SmallCNN.
\end{enumerate}

\subsection{Practical Implications}

These findings and learnings directly translate into several practical recommendations for deep learning practitioners implementing binary classification systems or multi-label classification systems:

\subsubsection{Guidelines for Output Layer Selection}

\begin{enumerate}
\item \textbf{Default Choice}: Based on our results and findings single-neuron sigmoid approach should be the default choice for binary classification tasks particularly for smaller or custom architectures where every parameter matters and model is solely based on less advanced architectures of neural networks.

\item \textbf{Model Size Considerations}: 
   \begin{itemize}
   \item For small to medium-sized models single-neuron approach offers notable performance benefits and go to solution.
   \item For very large pre-trained models (like ResNet50 or ViT) either approach will likely yield good results though the single-neuron approach still maintains a small edge over other approach.
   \end{itemize}

\item \textbf{Training Data Volume}: The advantage of the single-neuron approach becomes more pronounced with limited training data making it especially suitable for domains where labeled data is scarce or expensive to obtain or quality of data is not good because single neuron approach becomes safer choice.
\end{enumerate}

\subsubsection{Implementation Recommendations}

\begin{enumerate}
\item \textbf{Transfer Learning}: When fine-tuning pre-trained models for binary classification replacing the original output layer of the model with a single-neuron design yields optimal results even when the original model was trained with softmax outputs.

\item \textbf{Early Stopping Strategy}: Models with single-neuron outputs usually tend to converge faster so early stopping criteria may need adjustment compared to dual-neuron counterparts if dual neuron model is used. From our learnings and findings we recommend monitoring validation loss with a patience of 5-10 epochs because it suits well.

\item \textbf{Learning Rate Scheduling}: Single-neuron models usually benefit from a slightly higher initial learning rate while dual-neuron models may require more conservative learning rates to prevent instability.
\end{enumerate}

\subsubsection{Special Considerations}

\begin{enumerate}
\item \textbf{Model Scaling}: When scaling from binary to multi-class problems practitioners often choose dual-neuron outputs for binary problems to maintain architectural consistency specially in multiclass frameworks. Our results suggest this consistency comes at a small but measurable performance cost.

\item \textbf{Probability Calibration}: If well calibrated probability estimates are critical for the application (e.g in risk assessment systems) the single-neuron approach tends to provide better calibrated probabilities out-of-the-box though calibration techniques can improve both approaches.

\item \textbf{Deployment Constraints}: The single-neuron approach has a marginally smaller memory footprint which is result of less parameters and less calculations for weights update which may be beneficial in resource-constrained deployment environments like mobile devices.
\end{enumerate}